---
title: "jensen_shannon_stuff"
author: "githubz0r"
date: "2019-05-17"
output: workflowr::wflow_html
---

## Introduction

```{r}
library(conos)
library(pheatmap)
#library(fuck)
source('/home/larsc/SecretUtils/R/asdf.R')
source('/home/larsc/SecretUtils/R/peter_code_utils.R')
require(pagoda2)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(irlba)
con_object <- readRDS(file.path('/home/larsc/data/10x_preproced_graphed.rds'))
annot <- readRDS(file.path('/home/demharters/R/projects/UPF9_14_17_19_22_23_24_32_33/metadata_10x_final.rds'))
```

rowbind the adjusted expression values 
```{r}
if (is.null(annot$cellid)) {
  annot$cellid <- annot %>% rownames
}
annot <- annot %>% mutate(subtype_condition = paste(annot$subtype, annot$condition, sep='_'))
rbound_panel <- RbindPanel(con_object)
```

Are the cell names in the same order?
```{r}
identical(rownames(rbound_panel), annot$cellid) # thank god, maybe add a sort for the future in Panelize
```


```{r}
state_split <- split(annot, annot$condition, drop=TRUE)
condition_tables <- state_split %>% lapply(function(x){table(x$subtype)})
condition_tables$epilepsy<condition_tables$healthy # only Lamp5
condition_tables
```

PCA for later use 
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 3000)
pca_cm <- prcomp_irlba(rbound_panel[, od_genes],n=100)
#pca_cmat <- pca_cm$x[, 1:100]
pca_cmat <- pca_cm$x
rownames(pca_cmat) <- rownames(rbound_panel)
```


Bhattacharyya
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
healthy_probs <- subtype_split$healthy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=0)
epilepsy_probs <- subtype_split$epilepsy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=0)

Bhattacharyya <- function(x,y){
  -log(sum(sqrt(x*y)))
}
all_bhats <- Map(Bhattacharyya, healthy_probs, epilepsy_probs) %>% as_tibble
all_bhats <- gather(all_bhats, key=subtype, value=bhat_div)
ggplot(all_bhats, aes(y=bhat_div, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Cell by Cell Bhat, 100 cells each, 1000 od genes
```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
healthcellprobs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, od_genes)
epscellprobs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, od_genes)
all_singlecell_dists <- Map(CalculateAllBhat, healthcellprobs, epscellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=bhat)
ggplot(all_scd_gathered, aes(y=bhat, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) #+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)
```

Dependency for single subtypes
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 3000)
nr_cells <- c(30, 60, seq(100, 1000, by=50))
nr_cells2 <- (seq(30,300,by=15))
do_dist_cells <- function(nr.cell, subtype, pseudo.prob=1e-8, dist=js) {
  healthy_probs <- subtype_split$healthy[[subtype]] %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=nr.cell, pseudo.count=pseudo.prob)
  epilepsy_probs <- subtype_split$epilepsy[[subtype]] %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=nr.cell, pseudo.count=pseudo.prob)
  return(Bhattacharyya(healthy_probs, epilepsy_probs))
}

get_all_jscells <- function(subtype, cell.nr.vec, pseudo.prob=1e-8, dist='correlation'){
  js_nrcell <- cell.nr.vec %>% lapply(do_dist_cells, subtype, pseudo.prob, dist)
  jscell_df <- bind_cols(value=unlist(js_nrcell), nr.cells=cell.nr.vec) %>% mutate(subtype=subtype)
  #return(ggplot(jscell_df, aes(y=value, x=nr.cells))+geom_point()+ggtitle(subtype))
}

#all_jscells <- annot$subtype %>% unique %>% lapply(get_all_jscells, nr_cells)

do_plot <- function(sub.df){
  p <- ggplot(sub.df, aes(x=nr.cells, y=value, col=subtype))+geom_line()+
    theme(axis.text.x = element_text(angle = -90, hjust = 1))
  return(p)
}

#all_plots <- all_jscells %>% lapply(do_plot)
#subl1 <- do.call(rbind, all_jscells[1:5])

#p <- ggplot(all_jscells[[20]], aes(x=nr.cells, y=value, col=subtype))+geom_point()+theme(axis.text.x = element_text(angle = -90, hjust = 1))
#plot_grid(plotlist=all_plots[1:4], nrow=2)
subtypes_ <- annot$subtype %>% unique
#sub_jscells <- get_all_jscells('L2_Lamp5', nr_cells2, dist='correlation')
sub_jscells_rep <- replicate(3, list(get_all_jscells('L2_Lamp5', nr_cells2, dist='correlation'))) %>% bind_rows
p <- sub_jscells_rep %>% ggplot(aes(x=nr.cells, y=value, col=subtype))+geom_point()+theme(axis.text.x = element_text(angle = -90, hjust = 1))
p

```

try JBLD
```{r}
GetSubMatsList <- function(list.of.annots, rbound.panel, list.of.genes, cellid.col) {
  GetMat <- function(sub.annot, genes, rbound.panel, cellid.col) {
    cellids <- sub.annot[, cellid.col]
    return(rbound.panel[cellids, genes])
  }
  exps.list <- mapply(GetMat, list.of.annots, list.of.genes, MoreArgs=list(rbound.panel, cellid.col), SIMPLIFY=F)
  return(exps.list)
}

od_genes = conos:::getOdGenesUniformly(con_object$samples, 2000)
# listify
od_genes_list <- rep(list(od_genes), 20)
pca_genes_list <- rep(list(colnames(pca_cmat)), 20)

# split
state_split <- split(annot, annot$condition, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})

#cmat <- rbound_panel
state_sub_mats <- subtype_split %>% lapply(GetSubMatsList, pca_cmat, pca_genes_list, 4)

health_covs <- state_sub_mats$healthy %>% lapply(function(x){return(x %>% as.matrix %>% cov)})
epilepsy_covs <- state_sub_mats$epilepsy %>% lapply(function(x){return(x %>% as.matrix %>% cov)})

CalculateJBLD <- function(cov1, cov2) {
  JBLD <- log(det((cov1+cov2)/2)) -0.5*log(det(cov1%*%cov2))
  return(JBLD)
}

det(health_covs$Id2_Nos1%*%epilepsy_covs$Id2_Nos1)
all_jblds <- Map(CalculateJBLD, health_covs, epilepsy_covs)
jblds_gathered <- gather(all_jblds %>% as.data.frame, key=subtype, value=jbld)
ggplot(jblds_gathered, aes(y=jbld, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Correlation distance metric
fuck, try correlation matrix distance
```{r}
cmat <- rbound_panel
state_split <- split(annot, annot$condition, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
state_sub_mats <- subtype_split %>% lapply(GetSubMatsList, pca_cmat, pca_genes_list, 4)

health_cors <- state_sub_mats$healthy %>% lapply(cor)
epilepsy_cors <- state_sub_mats$epilepsy %>% lapply(cor)
all_cmds <- Map(CMD, health_cors, epilepsy_cors)
cmds_gathered <- gather(all_cmds %>% as.data.frame, key=subtype, value=cmd)
ggplot(cmds_gathered, aes(y=cmd, x=subtype)) +geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

P-values
load my de genes & get some DE genes, equal numbers for each subtype
```{r}
de_genez <- readRDS(file.path('/home/larsc/data/eps_10x_de_celltype.rds'))
```

```{r}
GetTopGenes <- function(de.result, nr, value='gene.names', filter.mito=TRUE){
  res.df <- de.result$res
  res.df <- res.df %>% dplyr::mutate(gene.names = rownames(res.df))
  if (filter.mito){
    res.df <- dplyr::filter(res.df, !stringr::str_detect(gene.names, "^MT-"))
  }
  res.sorted <- res.df %>% arrange(padj)
  return(res.sorted[[value]][1:nr])
}

top_genez <- de_genez %>% lapply(GetTopGenes, 100, filter.mito=FALSE)

# check how many DE genes for each
countPadjes <- function(de.result,threshold=0.1){
  padjes <- de.result$res$padj
  sum(na.omit(padjes<threshold))
}
subtypes_total_count <- split(annot$subtype, annot$subtype) %>% lapply(length)
padjes_count <- de_genez %>% lapply(countPadjes,threshold=0.05) %>% as.data.frame %>% gather(key=subtype, value=count) %>% 
  mutate(sumcells = as.numeric(table(annot$subtype)))
padjes_count %>% ggplot(aes(x=subtype, y=count))+geom_bar(stat='identity')+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Counts as function of number of cells
```{r}
padjes_count %>% ggplot(aes(x=sumcells, y=count))+geom_point()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```


Distribution of p values for each subtype
```{r}
top_padjes <- de_genez %>% lapply(GetTopGenes, 1000, value='padj', filter.mito=FALSE)
top_padjes_tibble <- top_padjes %>% as_tibble %>% gather(key=subtype, value=val) %>% mutate(log_val=log(val))
top_avg <- top_padjes %>% lapply(mean) %>% as.data.frame %>% gather(key=subtype, value=val) %>% mutate(log_val=log(val)) %>% 
  mutate(sumcells = as.numeric(table(annot$subtype)))

top_padjes_tibble %>% ggplot(aes(x=subtype, y=-log(val)))+geom_jitter()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

```{r}
top_avg %>% ggplot(aes(x=subtype, y=-log_val))+#geom_bar(stat='identity')+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))+ geom_point()+ggtitle('average pvalue using top 1000 DE genes')
```

As a function of number of points
```{r}
top_avg %>% ggplot(aes(x=sumcells, y=-log_val))+#geom_bar(stat='identity')+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))+ geom_point()+ggtitle('average pvalue using top 1000 DE genes')
```

