---
title: "pancreas_indrop"
author: "githubz0r"
date: "2019-16-6"
output: workflowr::wflow_html
---

load packages

```{r}
library(conos)
library(tidyverse)
devtools::load_all('/home/larsc/SecretUtils')
require(pagoda2)
library(pheatmap)
library(irlba)
library(igraph)
con_pancreas <- readRDS('/home/larsc/data/pancreas_indrop_conos_precced_graphed.rds')
location_of_annot <- '/d0-mendel/home/viktor_petukhov/SmallProjects/conos/data/conditions/seurat_islets/indrop/design_info.csv'
pancreas_annot <- location_of_annot %>% data.table::fread(sep=",",header=T) %>% as.data.frame() 
```


Rbind panels from conos objects
```{r}
rbound_pancreas <- RbindPanel(con_pancreas)
all.equal(rownames(rbound_pancreas), pancreas_annot$CellId)
```

Make groups for plotting
```{r}
pancreas_annot$subtype_disease <- paste0(pancreas_annot$CellType, pancreas_annot$Disease)

individual_annot <- setNames(pancreas_annot$Individual, pancreas_annot$CellId)
disease_annot <- setNames(pancreas_annot$Disease, pancreas_annot$CellId)
celltype_annot <- setNames(pancreas_annot$CellType, pancreas_annot$CellId)
sub_cond_annot <- setNames(pancreas_annot$subtype_disease, pancreas_annot$CellId)
```

```{r}
table(individual_annot)
table(celltype_annot)
```

Plot graph with different annotations
```{r}
con_pancreas$plotGraph(groups=disease_annot, font.size=3, size=0.3, alpha=0.3, show.legend=T)
con_pancreas$plotGraph(groups=celltype_annot, font.size=3, size=0.3, alpha=0.3, show.legend=T)
con_pancreas$plotGraph(groups=sub_cond_annot, font.size=3, size=0.3, alpha=0.3)
con_pancreas$plotGraph(groups=individual_annot, font.size=3, size=0.3, alpha=0.3, show.legend=T)
```

Initiate some variables
```{r}
od_genes = conos:::getOdGenesUniformly(con_pancreas$samples, 3000)
state_split <- split(pancreas_annot, pancreas_annot$Disease, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$CellType, drop=TRUE)})
```


Jensen Shannon, overall (microglia has by far the most cells so this will heavily skew the result due to dropout)
```{r}
he_probs <- subtype_split$normal %>% GetSampProbs(rbound_pancreas, od_genes, cellid.col = 1, pseudo.count=10^(-8))
t2_probs <- subtype_split$`type II diabetes mellitus` %>% 
  GetSampProbs(rbound_pancreas, od_genes, cellid.col = 1, pseudo.count=10^(-8))

all_dists <- Map(JensenShannon, he_probs, t2_probs) %>% as_tibble
all_dists_gathered <- gather(all_dists, key=subtype, value=js_distance)
ggplot(all_dists_gathered, aes(y=js_distance, x=subtype)) +geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```


Violins without neighbour smoothing?
```{r}
hecellprobs <- IndividualCellProbs(state_split$normal, rbound_pancreas, 3, 2, 100, od_genes, 10^(-8))
t2cellprobs <- IndividualCellProbs(state_split$`type II diabetes mellitus`, rbound_pancreas, 3, 2, 100, od_genes, 10^(-8))
all_singlecell_dists <- Map(CalculateAllJSD, hecellprobs, t2cellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=jsd)
ggplot(all_scd_gathered, aes(y=jsd, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```


PCA for correlation
```{r}
pca_cm <- prcomp_irlba(rbound_pancreas[, od_genes],n=100)
pca_cmat <- pca_cm$x
rownames(pca_cmat) <- rownames(rbound_pancreas)
pca_genes <- colnames(pca_cmat)
```

```{r}
he_vecs <- subtype_split$normal %>% GetSubMatrices(pca_cmat, pca_genes, cellid.col = 3, avg=T)
t2_vecs <- subtype_split$`type II diabetes mellitus` %>% GetSubMatrices(pca_cmat, pca_genes, cellid.col = 3, avg=T)

#t2_vecs <- subtype_split$AD %>% GetSampProbs(pca_cmat, pca_genes, cellid.col = 1, pseudo.count=0) # remember sign
#he_vecs <- subtype_split$WT %>% GetSampProbs(pca_cmat, pca_genes, cellid.col = 1, pseudo.count=0)

all_dists <- Map(function(x,y){1-cor(x,y)}, he_vecs, t2_vecs) %>% as_tibble
all_dists_gathered <- gather(all_dists, key=subtype, value=corcomplement)
ggplot(all_dists_gathered, aes(y=corcomplement, x=subtype)) +geom_bar(stat='identity') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```
Fractional plot
```{r}
FractionalPlot(pancreas_annot$Individual, pancreas_annot$CellType, pancreas_annot$Disease)
```

PAGA
```{r}
pancreas_adj <- igraph::as_adj(con_pancreas$graph, attr="weight")[pancreas_annot$CellId, pancreas_annot$CellId]

subtype_order <- (paste0(pancreas_annot$CellType) %>% unique)[order(paste0(pancreas_annot$CellType) %>% unique)]
membership_vec <- as.numeric(factor(pancreas_annot$subtype_disease))

pancreas_annot$subtype_sample <- paste0(pancreas_annot$CellType, '-', pancreas_annot$Individual)
membership_levels_subsamp <- factor(pancreas_annot$subtype_sample) %>% levels
membership_vec_subsamp <- as.numeric(factor(pancreas_annot$subtype_sample))
```


```{r}
connectivities <- GetPagaMatrix(pancreas_adj, membership_vec, scale=F)
linearized_stats <- seq(1, dim(connectivities)[1], 2) %>% sapply(function(i){connectivities[i,i+1]})

paga_df <- bind_cols(value=linearized_stats, subtype=subtype_order)
ggplot(paga_df, aes(y=-linearized_stats, x=subtype)) +geom_point()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

test with sample information
```{r}
connectivities <- GetPagaMatrix(pancreas_adj, membership_vec_subsamp, scale=F)
plot_df <- GeneratePagaSubSampDF(connectivities, pancreas_annot$CellType, pancreas_annot$Individual, pancreas_annot$Disease)

plot_df %>% ggplot(aes(x=subtype, y=-value ,dodge=condition,fill=condition))+
    geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                           axis.text.y = element_text(angle = 90, hjust = 0.5))+
  theme(legend.position="top")
```


Unaligned graph
```{r}
devtools::load_all('/home/larsc/SecretUtils')
raw_pancreas <- RbindRaw(con_pancreas)
pancreas_unaligned_adj <- GenerateUnalignedAdj(raw_pancreas, 
                                            cellid.vector=pancreas_annot$CellId)

```

```{r}
connectivities <- GetPagaMatrix(pancreas_unaligned_adj, membership_vec, scale=F)
linearized_stats <- seq(1, dim(connectivities)[1], 2) %>% sapply(function(i){connectivities[i,i+1]})

paga_df <- bind_cols(value=linearized_stats, subtype=subtype_order)
ggplot(paga_df, aes(y=-linearized_stats, x=subtype)) +geom_point()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```


