---
title: "simulations_plots"
author: "githubz0r"
date: "2019-07-12"
output: workflowr::wflow_html
---

```{r}
library(conos)
library(tidyverse)
devtools::load_all('/home/larsc/SecretUtils')
library(cowplot)
library(splatter)
devtools::load_all('/home/viktor_petukhov/Copenhagen/NeuronalMaturation')
```
Load data

```{r, message=FALSE}
gene_names <- paste0('Gene', '', 1:10000)
gene_names2 <- paste0('Gene', '', 1:20000)

cellcm1 <- readMM('/home/larsc/data/splatter_data/cell1cm.mtx') %>% as('dgCMatrix')
cellcm2 <- readMM('/home/larsc/data/splatter_data/cell2cm.mtx') %>% as('dgCMatrix')
cellcm3 <- readMM('/home/larsc/data/splatter_data/cell3cm.mtx') %>% as('dgCMatrix')
cellannot1 <- read_csv('/home/larsc/data/splatter_data/cell1annot.csv')
cellannot2 <- read_csv('/home/larsc/data/splatter_data/cell2annot.csv')
cellannot3 <- read_csv('/home/larsc/data/splatter_data/cell3annot.csv')
rownames(cellcm1) <- cellannot1$cellid; colnames(cellcm1) <- gene_names
rownames(cellcm2) <- cellannot2$cellid; colnames(cellcm2) <- gene_names
rownames(cellcm3) <- cellannot3$cellid; colnames(cellcm3) <- gene_names

covercm1 <- readMM('/home/larsc/data/splatter_data/cover1cm.mtx') %>% as('dgCMatrix')
covercm2 <- readMM('/home/larsc/data/splatter_data/cover2cm.mtx') %>% as('dgCMatrix')
covercm3 <- readMM('/home/larsc/data/splatter_data/cover3cm.mtx') %>% as('dgCMatrix')
coverannot1 <- read_csv('/home/larsc/data/splatter_data/cover1annot.csv')
coverannot2 <- read_csv('/home/larsc/data/splatter_data/cover2annot.csv')
coverannot3 <- read_csv('/home/larsc/data/splatter_data/cover3annot.csv')
rownames(covercm1) <- coverannot1$cellid; colnames(covercm1) <- gene_names
rownames(covercm2) <- coverannot2$cellid; colnames(covercm2) <- gene_names
rownames(covercm3) <- coverannot3$cellid; colnames(covercm3) <- gene_names

genecm1 <- readMM('/home/larsc/data/splatter_data/gene1cm.mtx') %>% as('dgCMatrix')
genecm2 <- readMM('/home/larsc/data/splatter_data/gene2cm.mtx') %>% as('dgCMatrix')
genecm3 <- readMM('/home/larsc/data/splatter_data/gene3cm.mtx') %>% as('dgCMatrix')
geneannot1 <- read_csv('/home/larsc/data/splatter_data/gene1annot.csv')
geneannot2 <- read_csv('/home/larsc/data/splatter_data/gene2annot.csv')
geneannot3 <- read_csv('/home/larsc/data/splatter_data/gene3annot.csv')
rownames(genecm1) <- geneannot1$cellid; colnames(genecm1) <- gene_names2
rownames(genecm2) <- geneannot2$cellid; colnames(genecm2) <- gene_names2
rownames(genecm3) <- geneannot3$cellid; colnames(genecm3) <- gene_names2
```

First make pagoda objects with reductions and graphs
```{r, message=FALSE}
cell1p2 <- NeuronalMaturation::GetPagoda(Matrix::t(cellcm1), n.odgenes=3000, embeding.type=NULL)
cell2p2 <- NeuronalMaturation::GetPagoda(Matrix::t(cellcm2), n.odgenes=3000, embeding.type=NULL)
cell3p2 <- NeuronalMaturation::GetPagoda(Matrix::t(cellcm3), n.odgenes=3000, embeding.type=NULL)
saveRDS(cell1p2, '/home/larsc/data/splatter_data/cell1p2.rds')
saveRDS(cell2p2, '/home/larsc/data/splatter_data/cell2p2.rds')
saveRDS(cell3p2, '/home/larsc/data/splatter_data/cell3p2.rds')

cover1p2 <- NeuronalMaturation::GetPagoda(Matrix::t(covercm1), n.odgenes=3000, embeding.type=NULL)
cover2p2 <- NeuronalMaturation::GetPagoda(Matrix::t(covercm2), n.odgenes=3000, embeding.type=NULL)
cover3p2 <- NeuronalMaturation::GetPagoda(Matrix::t(covercm3), n.odgenes=3000, embeding.type=NULL)
saveRDS(cover1p2, '/home/larsc/data/splatter_data/cover1p2.rds')
saveRDS(cover2p2, '/home/larsc/data/splatter_data/cover2p2.rds')
saveRDS(cover3p2, '/home/larsc/data/splatter_data/cover3p2.rds')

gene1p2 <- NeuronalMaturation::GetPagoda(Matrix::t(genecm1), n.odgenes=3000, embeding.type=NULL)
gene2p2 <- NeuronalMaturation::GetPagoda(Matrix::t(genecm2), n.odgenes=3000, embeding.type=NULL)
gene3p2 <- NeuronalMaturation::GetPagoda(Matrix::t(genecm3), n.odgenes=3000, embeding.type=NULL)
saveRDS(gene1p2, '/home/larsc/data/splatter_data/gene1p2.rds')
saveRDS(gene2p2, '/home/larsc/data/splatter_data/gene2p2.rds')
saveRDS(gene3p2, '/home/larsc/data/splatter_data/gene3p2.rds')
```


```{r}
cellpaga1 <- MakeSimRepPaga(cellannot1, cell1p2, varied.factor='ncells')
cellpaga2 <- MakeSimRepPaga(cellannot2, cell2p2, varied.factor='ncells')
cellpaga3 <- MakeSimRepPaga(cellannot3, cell3p2, varied.factor='ncells')
bound_cellpaga <- bind_rows(cellpaga1, cellpaga2, cellpaga3)
#bound_cellpaga <- readRDS('/home/larsc/data/simcellpagadf.rds')
bound_cellpaga %>% ggplot(aes(x=ncell, y=paga.connectivity.value, col=de.prob))+
  geom_jitter(size=1, alpha=0.9)
cellpaga2 %>% filter(ncell==500) %>% ggplot(aes(x=ncell, y=paga.connectivity.value, col=de.prob))+geom_point(size=1, alpha=0.9)
```

the mean shape vector associated with the lib.loc vector is meanshapevec <- c(0.1, 0.2, 0.3, 0.4, 0.5)
```{r}
coverpaga1 <- MakeSimRepPaga(coverannot1, cover1p2, varied.factor='cover')
coverpaga2 <- MakeSimRepPaga(coverannot2, cover2p2, varied.factor='cover')
coverpaga3 <- MakeSimRepPaga(coverannot3, cover3p2, varied.factor='cover')
bound_coverpaga <- bind_rows(coverpaga1, coverpaga2, coverpaga3)

bound_coverpaga %>% ggplot(aes(x=lib.loc, y=paga.connectivity.value, col=de.prob))+geom_jitter(size=1, alpha=0.9)
coverpaga1 %>% filter(lib.loc==8) %>% ggplot(aes(x=lib.loc, y=paga.connectivity.value, col=de.prob))+
  geom_point(size=1, alpha=0.9)
```
Something strange might be happening here. We select select lib.loc 8 of one of the replicates and perform this one separately.
```{r}
covertestannot <- (coverannot2 %>% split(coverannot2$lib.loc))$`8`
cover8test <- covercm2[covertestannot$cellid, ]
covertestpaga <- MakeSimRepPaga(covertestannot, cover8test, varied.factor='cover')
covertestpaga %>% ggplot(aes(x=lib.loc, y=paga.connectivity.value, col=de.prob))+geom_point() # looks correct
```



```{r}
genepaga1 <- MakeSimRepPaga(geneannot1, gene1p2, varied.factor='ngenes')
genepaga2 <- MakeSimRepPaga(geneannot2, gene2p2, varied.factor='ngenes')
genepaga3 <- MakeSimRepPaga(geneannot3, gene3p2, varied.factor='ngenes')
bound_genepaga <- bind_rows(genepaga1, genepaga2, genepaga3)

bound_genepaga %>% ggplot(aes(x=ngenes, y=paga.connectivity.value, col=de.prob))+geom_jitter(size=1, alpha=0.9)
genepaga1 %>% ggplot(aes(x=ngenes, y=paga.connectivity.value, col=de.prob))+geom_point(size=1, alpha=0.9)
```

Correlation
```{r}
cec1 <- cell1p2$reductions$PCA; pca_genes <- cec1 %>% colnames
cec2 <- cell2p2$reductions$PCA
cec3 <- cell3p2$reductions$PCA
coc1 <- cover1p2$reductions$PCA
coc2 <- cover2p2$reductions$PCA
coc3 <- cover3p2$reductions$PCA
gec1 <- gene1p2$reductions$PCA
gec2 <- gene2p2$reductions$PCA
gec3 <- gene3p2$reductions$PCA

cellmats <- list(cec1, cec2, cec3)
cellannots <- list(cellannot1, cellannot2, cellannot3)
covmats <- list(coc1, coc2, coc3)
covannots <- list(coverannot1, coverannot2, coverannot3)
genemats <- list(gec1, gec2, gec3)
geneannots <- list(geneannot1, geneannot2, geneannot3)

docorr <- function(a.mat, an.annot){
  corr <- performDistance(an.annot$subtype, an.annot$cellid, an.annot$condition, a.mat, pca_genes)
}
corrall <- function(mat.list, annot.list, factor.identity) {
  cors <- Map(docorr, mat.list, annot.list)
  cors.all <- dplyr::bind_rows(cors)
  if (factor.identity!='coverage'){
    cors.all %<>% mutate(varied.factor=as.factor(as.numeric(gsub(' .*', '', subtype))), 
                       de.prob=as.factor(as.numeric(gsub('.* ', '', subtype))))
    cors.all %<>% dplyr::rename(!!factor.identity:=varied.factor)
  } else {
    cors.all %<>% mutate(varied.factor=gsub(' .*', '', subtype), 
                       de.prob=as.factor(as.numeric(gsub('.* ', '', subtype))))
    cors.all %<>% dplyr::rename(!!factor.identity:=varied.factor)
  }
  return(cors.all)
}

cellcor <- corrall(cellmats, cellannots, factor.identity='ncell')
covcor <- corrall(covmats, covannots, factor.identity='coverage')
genecor <- corrall(genemats, geneannots, factor.identity='ngenes')

cellcor %>% ggplot(aes(x=ncell, y=correlation.distance, col=de.prob))+geom_point(size=1, alpha=0.9)
cellcor %>% filter(ncell==1000) %>% ggplot(aes(x=ncell, y=correlation.distance, col=de.prob))+geom_jitter(size=1, alpha=0.9)
covcor %>% ggplot(aes(x=coverage, y=correlation.distance, col=de.prob))+geom_jitter(size=1, alpha=0.9)
covcor %>% filter(coverage=='8_0.4') %>% ggplot(aes(x=coverage, y=correlation.distance, col=de.prob))+geom_point(size=1, alpha=0.9)
genecor %>% ggplot(aes(x=ngenes, y=correlation.distance, col=de.prob))+geom_point(size=1, alpha=0.9)
genecor %>% filter(ngenes==1000) %>% ggplot(aes(x=ngenes, y=correlation.distance, col=de.prob))+geom_jitter(size=1, alpha=0.9)
```

That lib.loc=8 coverage is still bothing me.
```{r}
covtestcmp2 <- cover8test %>% Matrix::t() %>% NeuronalMaturation::GetPagoda(n.odgenes=3000, embeding.type=NULL)
covtestpca <- covtestcmp2$reductions$PCA
testcorr <- corrall(list(covtestpca), list(covertestannot), 'coverage')
testcorr %>% ggplot(aes(x=coverage, y=correlation.distance, col=de.prob))+geom_point(size=1, alpha=0.9)
```

