---
title: "subtype_distances"
author: "githubz0r"
date: "2019-03-26"
output: workflowr::wflow_html

---

## Introduction

First we load packages and data (has been preproced with pagoda2 and graph builth with conos)
```{r}
library(conos)
library(pheatmap)
#library(fuck)
source('/home/larsc/SecretUtils/R/asdf.R')
require(pagoda2)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(cowplot)
con_object <- readRDS(file.path('/home/larsc/data/10x_preproced_graphed.rds'))
annot <- readRDS(file.path('/home/demharters/R/projects/UPF9_14_17_19_22_23_24_32_33/metadata_10x_final.rds'))
```

rowbind the adjusted expression values 
```{r}
rbound_panel <- RbindPanel(con_object)
```

Are the cell names in the same order?
```{r}
identical(rownames(rbound_panel), rownames(annot)) # thank god, maybe add a sort for the future in Panelize
```


plot joint graph
```{r}
con_object$plotGraph(color.by='sample',mark.groups=F,alpha=0.1,show.legend=T)
```

set cell annotation factor thingy
```{r}
annot <- annot %>% mutate(cellids = rownames(annot)) # turn rownames into a col for convenience
annot <- annot %>% mutate(subtype_condition = paste(annot$subtype, annot$condition, sep='_'))
cellannot=setNames(annot[,5], annot[,4])
```

Let's see the graph now
```{r}
con_object$plotGraph(groups=cellannot, font.size=3, shuffle.colors=T, show.legend=F)
```
Damn, it looks awful.

Check which samples have which subtypes
```{r}
# just for interest check how many subtypes for each sample
howmanysubs <- function(x, annotation){
  annotsubset <- annotation[annotation$sample==x,]
  return(unique(annotsubset$subtype)) #%>% length)
}
subs_of_samples <- unlist(CreateSampleGroups(annot, 3, 1)) %>% sapply(howmanysubs, annot)
# healthy 2 and 4 missing one
samp_intersect <- subs_of_samples %>% Reduce(intersect,.)
missing_subs <- setdiff(subs_of_samples$epilepsy1, samp_intersect)
missing_subs
```

Inter sample distance, aggregated
```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 5000)
# split annot in categories (disease, healthy)
state_split <- split(annot, annot$condition, drop=TRUE)
# get probability distributions for subtype/sample categories
all_probs <- ObtainProbabilities(state_split, rbound_panel, 1, 2, 4, od_genes, pseudo.count=10^(-6))
# now get JS distances
all_distances <- Map(CalculateAllJSD, all_probs$healthy, all_probs$epilepsy)
#all_dists_dfd <- all_distances %>% lapply(function(x){data.frame(comparison=names(x), value=x, row.names=NULL)})
# use stackexchange magic to pad NAs to shorter distance vectors, and also tibbling
all_dists <- lapply(all_distances, `length<-`, max(lengths(all_distances))) %>% as_tibble
all_dists <- all_dists %>% mutate(comparison=names(all_distances$Id2_Reln))
# gather into format for easy ggplot
all_dists_gathered <- gather(all_dists, key=subtype, value=js_distance, -comparison)
# plot
ggplot(all_dists_gathered, aes(y=js_distance, x=subtype)) + geom_violin()+geom_point(aes(col=comparison), alpha = 0.6, size=1.8) +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```



```{r}
disannot<-setNames(annot[,3], annot[,4])
subannot<-setNames(annot[,2], annot[,4])
con_object$plotGraph(groups=subannot, alpha=0.1, font.size=3)
```

```{r}
con_object$plotGraph(groups=disannot, alpha=0.1)
```



let's see overall JS dist.
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
state_split <- split(annot, annot$condition, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
healthy_probs <- subtype_split$healthy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=10^(-10))
epilepsy_probs <- subtype_split$epilepsy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=10^(-10))

all_dists <- Map(JensenShannon, healthy_probs, epilepsy_probs) %>% as_tibble
all_dists_gathered <- gather(all_dists, key=subtype, value=js_distance)
ggplot(all_dists_gathered, aes(y=js_distance, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```


Dependency on nr of cells or genes?
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 2000)
state_split <- split(annot, annot$condition, drop=TRUE)
condition_tables <- state_split %>% lapply(function(x){table(x$subtype)})
condition_df <- condition_tables %>% as.data.frame
condition_df <- condition_df %>% mutate(min_cells = mapply(min, epilepsy.Freq, healthy.Freq))
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
healthy_probs <- subtype_split$healthy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=10^(-10))
epilepsy_probs <- subtype_split$epilepsy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=10^(-10))

all_dists <- Map(JensenShannon, healthy_probs, epilepsy_probs) %>% as_tibble
all_dists_gathered <- gather(all_dists, key=subtype, value=js_distance) %>% mutate(min_cells = condition_df$min_cells)
ggplot(all_dists_gathered, aes(y=js_distance, x=min_cells, col=subtype)) +geom_point() +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 2000)
nr_cells <- c(30, 60, seq(100, 1000, by=50))
nr_cells2 <- (seq(30,300,by=15))
do_js_cells <- function(nr.cell, subtype, pseudo.prob=1e-8) {
  healthy_probs <- subtype_split$healthy[[subtype]] %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=nr.cell, pseudo.count=pseudo.prob)
  epilepsy_probs <- subtype_split$epilepsy[[subtype]] %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=nr.cell, pseudo.count=pseudo.prob)
  js.dist <- JensenShannon( healthy_probs, epilepsy_probs)
  return(js.dist)
}

get_all_jscells <- function(subtype, cell.nr.vec, pseudo.prob=1e-8){
  js_nrcell <- cell.nr.vec %>% lapply(do_js_cells, subtype)
  jscell_df <- bind_cols(value=unlist(js_nrcell), nr.cells=cell.nr.vec) %>% mutate(subtype=subtype)
  #return(ggplot(jscell_df, aes(y=value, x=nr.cells))+geom_point()+ggtitle(subtype))
}

all_jscells <- annot$subtype %>% unique %>% lapply(get_all_jscells, nr_cells)

do_plot <- function(sub.df){
  p <- ggplot(sub.df, aes(x=nr.cells, y=value, col=subtype))+geom_line()+
    theme(axis.text.x = element_text(angle = -90, hjust = 1))
  return(p)
}
all_plots <- all_jscells %>% lapply(do_plot)
subl1 <- do.call(rbind, all_jscells[1:5])
p <- ggplot(all_jscells[[20]], aes(x=nr.cells, y=value, col=subtype))+geom_point()+theme(axis.text.x = element_text(angle = -90, hjust = 1))
#plot_grid(plotlist=all_plots[1:4], nrow=2)
p

```
Number of genes?
```{r}
nr_genes=seq(100, 5000, 250)
do_js_genes <- function(nr.genes, pseudo.prob=1e-8) {
  od_genes = conos:::getOdGenesUniformly(con_object$samples, nr.genes)
  healthy_probs <- subtype_split$healthy$L2_Lamp5 %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=100, pseudo.count=pseudo.prob)
  epilepsy_probs <- subtype_split$epilepsy$L2_Lamp5 %>% 
    SelectCellProbsAggregated(od_genes, rbound_panel, cellid.col = 4, nr.cell=100, pseudo.count=pseudo.prob)
  js.dist <- JensenShannon( healthy_probs, epilepsy_probs)
  return(js.dist)
}
js_nrgenes <- nr_genes %>% lapply(do_js_genes)
jsgenes_df <- bind_cols(value=unlist(js_nrgenes), nr.genes=nr_genes)
ggplot(jsgenes_df, aes(y=value, x=nr.genes))+geom_point()+ggtitle('JSD for L2_Lamp5, using x number of genes from each condition')
```


load my de genes & get some DE genes, equal numbers for each subtype
```{r}
de_genez <- readRDS(file.path('/home/larsc/data/eps_10x_de_celltype.rds'))

GetTopGenes <- function(de.result, nr, value='gene.names', filter.mito=TRUE){
  res.df <- de.result$res
  res.df <- res.df %>% dplyr::mutate(gene.names = rownames(res.df))
  if (filter.mito){
    res.df <- dplyr::filter(res.df, !stringr::str_detect(gene.names, "^MT-"))
  }
  res.sorted <- res.df %>% arrange(padj)
  return(res.sorted[[value]][1:nr])
}

top_genez <- de_genez %>% lapply(GetTopGenes, 100, filter.mito=FALSE)

# check how many DE genes for each
countPadjes <- function(de.result,threshold=0.1){
  padjes <- de.result$res$padj
  sum(na.omit(padjes<threshold))
}
padjes_count <- de_genez %>% lapply(countPadjes,threshold=0.3) %>% as.data.frame %>% gather(key=subtype, value=count)
padjes_count %>% ggplot(aes(x=subtype, y=count))+geom_bar(stat='identity')+theme(axis.text.x = element_text(angle = -90, hjust = 1))

top_padjes <- de_genez %>% lapply(GetTopGenes, 1000, value='padj', filter.mito=FALSE)
top_avg <- top_padjes %>% lapply(mean) %>% as.data.frame %>% gather(key=subtype, value=val)
top_padjes %>% as_tibble %>% gather(key=subtype, value=val) %>% ggplot(aes(x=subtype, y=val))+geom_boxplot()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
top_avg %>% as_tibble %>% gather(key=subtype, value=val) %>% mutate(log_val=log(val)) %>% 
  ggplot(aes(x=subtype, y=log_val))+#geom_bar(stat='identity')+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))+ geom_point()

```

Just curious what the distributions of padjes looks like
```{r, fig.height=7, fig.width=10}
de_genes_df <- de_genez %>% lapply(GetTopGenes, 100, value='padj', filter.mito=FALSE) %>% as_tibble
de_genes_gather <- gather(de_genes_df, key=subtype, value=padj)
ggplot(de_genes_gather, aes(y=padj, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))#+geom_jitter(aes(col=subtype), alpha = 0.3, size=0.5)
```

top 100 de genes for each subtype individually
```{r, fig.height=7, fig.width=10}
ObtainDistributions <- function(sub.annot, some.genes, rbound.panel, cellid.col, nr.cell, pseudo.count){
    return(SelectCellProbs(sub.annot, rbound.panel, cellid.col, nr.cell, some.genes, pseudo.count))
  } # just to rearrange the inputs for mapply

IndividualCellProbsList <- function(annotation, rbound.panel, cellid.col, sub.col, nr.cell, genes.list, pseudo.count=0){
  sub.split <- split(annotation, annotation[, sub.col], drop=T)
  cell.distributions <- mapply(ObtainDistributions, sub.split, genes.list, MoreArgs=list(rbound.panel, cellid.col, nr.cell, pseudo.count), SIMPLIFY=F)
  return(cell.distributions)
}
ProbsToJSD <- function(state.split, rbound.panel, cellid.col, sub.col, nr.cell, genes.list, pseudo.count){
  list.of.lists <- state.split %>% lapply(IndividualCellProbsList, rbound.panel, cellid.col, sub.col, 
                                          nr.cell, genes.list, pseudo.count)
  all.sc.dists.de <- Map(CalculateAllJSD, list.of.lists$healthy, list.of.lists$epilepsy)
  return(all.sc.dists.de %>% as_tibble)
} # specific for this data but saves some lines

### some sanity checking, do not mind it
#test_splits <- state_split %>% lapply(function(x){split(x, x[,2], drop=T)})
#min_cells_eps <- test_splits$epilepsy %>% lapply(function(x){dim(x)[1]}) %>% unlist %>% min
#min_cells_healthy <- test_splits$healthy %>% lapply(function(x){dim(x)[1]}) %>% unlist %>% min
#test_nos1_probs<-test_splits$epilepsy$Id2_Nos1 %>% SelectCellProbs(rbound_panel, 4, 101, top_genez$Id2_Nos1, 10^(-8))
#test_cell <- names(test_nos1_probs)[1]
#test_mat <- rbound_panel[test_cell, top_genez$Id2_Nos1]
#test_dist <- test_mat/sum(test_mat)
#test_dist <- test_dist +10^(-8)
#test_dist <- test_dist/sum(test_dist)
#all.equal(test_dist, test_nos1_probs[[1]]) # tru
### end sanity check

# equal numbers
all_sc_dists_de <- ProbsToJSD(state_split, rbound_panel, 4, 2, 100, top_genez, 10^(-8))
all_scd_de_gathered <- gather(all_sc_dists_de, key=subtype, value=JSD)
ggplot(all_scd_de_gathered, aes(y=JSD, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))#+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)
```

Sam's genes, not equal number of genes
```{r, fig.height=7, fig.width=10}
de_genes <- readRDS("/d0-mendel/home/demharters/R/projects/UPF9_14_17_19_22_23_24_32_33/de_multilevel_genes.rds")
sam_de_count <- de_genes %>% lapply(length) %>% as.data.frame
#names(de_genez)==names(de_genes) # just checking
CheckOverlap <- function(x,y){ # is it completely off or wut
  y=y[1:100]
  return(sum(x %in% y)/100)
}
check_overlap <- Map(CheckOverlap, top_genez, de_genes) # low for pvalb nos since it only has 16
all_sc_dists_de_sam <- ProbsToJSD(state_split, rbound_panel, 4, 2, 100, de_genes, 10^(-8))
all_scd_de_gathered_sam <- gather(all_sc_dists_de_sam, key=subtype, value=JSD)
ggplot(all_scd_de_gathered_sam, aes(y=JSD, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))#+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)

```


Constant number of OD genes, constant number of cells.

```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
healthcellprobs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, od_genes, 10^(-8))
epscellprobs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, od_genes, 10^(-8))
all_singlecell_dists <- Map(CalculateAllJSD, healthcellprobs, epscellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=jsd)
ggplot(all_scd_gathered, aes(y=jsd, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) #+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)
```
WHat if we use cor instead?

Correlation (overall)
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 5000)
top_genez <- de_genez %>% lapply(GetTopGenes, 5000, filter.mito=FALSE)

state_split <- split(annot, annot$condition, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
healthy_avg <- subtype_split$healthy %>% GetSubMatrices(rbound_panel, od_genes, cellid.col = 4) %>% lapply(GetColMeans)
epilepsy_avg <- subtype_split$epilepsy %>% GetSubMatrices(rbound_panel, od_genes, cellid.col = 4) %>% lapply(GetColMeans)

#healthy_avg <- subtype_split$healthy %>% GetSubMatsList(rbound_panel, top_genez, cellid.col = 4) %>% lapply(GetColMeans)
#epilepsy_avg <- subtype_split$epilepsy %>% GetSubMatsList(rbound_panel, top_genez, cellid.col = 4) %>% lapply(GetColMeans)

all_dists <- Map(cor, healthy_avg, epilepsy_avg) %>% as_tibble
all_dists_gathered <- gather(all_dists, key=subtype, value=correlation)
ggplot(all_dists_gathered, aes(y=correlation, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
ggplot(all_dists_gathered, aes(y=correlation, x=subtype)) +geom_point() +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))+ylim(0.75,1)
```

pairwise correlation
```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 2000)
healthcellvecs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, od_genes, vec=T)
epscellvecs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, od_genes, vec=T)
all_singlecell_dists <- Map(CalculateAllCor, healthcellvecs, epscellvecs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=correlation)
ggplot(all_scd_gathered, aes(y=correlation, x=subtype)) + geom_boxplot()+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```



Constant number of OD genes, constant number of cells, normalize variance.
```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 100)
healthcellprobs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, od_genes, 10^(-8), norm.var=T)
epscellprobs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, od_genes, 10^(-8), norm.var=T)
all_singlecell_dists <- Map(CalculateAllJSD, healthcellprobs, epscellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=jsd)
ggplot(all_scd_gathered, aes(y=jsd, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) #+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)
```

Instead of OD genes, union of most DE genes
```{r, fig.height=7, fig.width=10}
top_genes_union <- top_genez %>% unlist %>% unique
healthcellprobs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, top_genes_union, 10^(-8))
epscellprobs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, top_genes_union, 10^(-8))
all_singlecell_dists <- Map(CalculateAllJSD, healthcellprobs, epscellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=jsd)
ggplot(all_scd_gathered, aes(y=jsd, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Now Aggregation of cells, same number of cells, 2000 od genes
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 2000)
od_genes_list <- rep(list(od_genes), 20)

IndividualCellProbsAgg <- function(annotation, rbound.panel, cellid.col, sub.col, nr.cell, genes.list, pseudo.count=0){
  sub.split <- split(annotation, annotation[, sub.col], drop=T)
  cell.distributions <- mapply(SelectCellProbsAggregated, sub.split, genes.list, MoreArgs=list(rbound.panel, cellid.col, nr.cell, pseudo.count), SIMPLIFY=F)
  return(cell.distributions)
}
sub_probs_agg <- state_split %>% lapply(IndividualCellProbsAgg, rbound_panel, 4, 2, 200, od_genes_list, 10^(-8))
jsd_agg <- Map(JensenShannon, sub_probs_agg$healthy, sub_probs_agg$epilepsy) %>% as_tibble
jsd_agg_gathered <- gather(jsd_agg, key=subtype, value=js_distance_cell_agg)
ggplot(jsd_agg_gathered, aes(y=js_distance_cell_agg, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Sanity checking coz' this weird (using previous iteration of prev.cell with top_genez instead of od genes)
```{r}
# let's double check this bullshit (only meaningful if using 2224 cells so we sample all in eps lamp5)
#test_sub_split <- state_split %>% lapply(function(x){split(x, x[,2])})
#test_mat <- rbound_panel[test_sub_split$epilepsy$L2_Lamp5[,4],top_genez$L2_Lamp5]
#test_probs <- test_mat %>% GetColMeans
#test_probs <- test_probs/sum(test_probs)
#test_probs <- test_probs+1e-8 # feels better to add pseudo-prob than pseudo-variance-normalized-count
#test_probs <- test_probs/sum(test_probs)
#test_probs2 <- sub_probs_agg$epilepsy$L2_Lamp5
#all.equal(test_probs, test_probs2) # so the cells x genes matrix into prob dist is correct

# are cell names correct?
#test_cellnames <- IndividualCellProbsList(state_split$epilepsy, rbound_panel, 4, 2, 2224, top_genez, pseudo.count=1e-8)
#sum(test_cellnames$L2_Lamp5 %>% names %in% rownames(test_mat))==2224 # it's true, so function works
```


Bhattacharyya
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
state_split <- split(annot, annot$condition, drop=TRUE)
subtype_split <- state_split %>% lapply(function(x){split(x, x$subtype, drop=TRUE)})
healthy_probs <- subtype_split$healthy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=0)
epilepsy_probs <- subtype_split$epilepsy %>% GetSampProbs(rbound_panel, od_genes, cellid.col = 4, pseudo.count=0)

Bhattacharyya <- function(x,y){
  -log(sum(sqrt(x*y)))
}
all_bhats <- Map(Bhattacharyya, healthy_probs, epilepsy_probs) %>% as_tibble
all_bhats <- gather(all_bhats, key=subtype, value=bhat_div)
ggplot(all_bhats, aes(y=bhat_div, x=subtype)) +geom_bar(stat='identity', col='hotpink') +
  theme(axis.text.x = element_text(angle = -90, hjust = 1))
```

Cell by Cell Bhat, 100 each, 1000 od genes
```{r, fig.height=7, fig.width=10}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 1000)
healthcellprobs <- IndividualCellProbs(state_split$healthy, rbound_panel, 4, 2, 100, od_genes)
epscellprobs <- IndividualCellProbs(state_split$epilepsy, rbound_panel, 4, 2, 100, od_genes)
all_singlecell_dists <- Map(CalculateAllBhat, healthcellprobs, epscellprobs)
all_sc_dists <- all_singlecell_dists %>% as_tibble
all_scd_gathered <- gather(all_sc_dists, key=subtype, value=bhat)
ggplot(all_scd_gathered, aes(y=bhat, x=subtype)) + geom_violin(aes(col=subtype))+
  theme(axis.text.x = element_text(angle = -90, hjust = 1)) #+geom_jitter(aes(col=subtype), alpha = 0.05, size=0.5)
```

Individual subtype joint graph plot
```{r}
plotOneSubtype <- function(con.object, annotation, subtype, font.size=2, alpha=0.3, size=0.4){
  split.annot<-split(annotation, annotation$subtype)
  sub.annot <- split.annot[[subtype]]
  sub.annot <- sub.annot %>% mutate(sub.cond = paste(sub.annot$subtype, sub.annot$condition, sep='_'))
  sub.groups <- setNames(sub.annot$sub.cond, sub.annot$cellids)
  con.object$plotGraph(groups=sub.groups, font.size=font.size, alpha=alpha, size=size, mark.groups=T, plot.na=F)
}
#plotOneSubtype(con_object, annot, 'L2_Lamp5') # really should use repel, but I can't make it work
all_types <- annot$subtype %>% unique
all_types_plots <- all_types %>% lapply(function(x, con.obj, annotation){plotOneSubtype(con.obj, annotation, x)},
                                        con_object, annot)
plot_grid(plotlist=all_types_plots, nrow=4)

```

```{r}
plotOneSubtype(con_object, annot, 'L4_Rorb', font.size=2, alpha=0.5, size=1.5)
(annot$subtype=='Pvalb_Nos1' & annot$condition=='epilepsy') %>% sum
(annot$subtype=='Pvalb_Nos1' & annot$condition=='healthy') %>% sum
# actually let me just count all of them
subtype_cond_table <- paste(annot$condition, annot$subtype) %>% table
# maybe should try enforcing equal numbers of cells
```

Samples
```{r, warning=FALSE}
plotOneSubtypeSamples <- function(con.object, annotation, subtype, font.size=2, alpha=0.3, size=0.4, show.legend=F){
  split.annot<-split(annotation, annotation$subtype)
  sub.annot <- split.annot[[subtype]]
  sub.annot <- sub.annot %>% mutate(sub.cond = paste(sub.annot$subtype, sub.annot$sample, sep='_'))
  sub.groups <- setNames(sub.annot$sub.cond, sub.annot$cellids)
  con.object$plotGraph(groups=sub.groups, font.size=font.size, alpha=alpha, size=size, mark.groups=T, plot.na=F, show.legend=show.legend)
}
plotOneSubtypeSamples(con_object, annot, 'L2_Lamp5', font.size=c(2,4), show.legend=T)
all_types_plots_samples <- all_types %>% lapply(function(x, con.obj, annotation){
  plotOneSubtypeSamples(con.obj, font.size=1.5, annotation, x)
  }, con_object, annot)
plot_grid(plotlist=all_types_plots_samples, nrow=4)
```



