---
title: "transmute_peters_code"
output: workflowr::wflow_html
---
load packages and data
```{r}
library(conos)
library(parallel)
library(ggplot2)
library(Matrix)
library(data.table)
library(pagoda2)
library(cowplot)
library(dplyr)
library(abind)
library(tidyr)
require(Rtsne)

# load data
scon <- readRDS("/d0-mendel/home/meisl/Workplace/BMME/a.data/Selected_Joint_embdding/conA.rds")
scon <- Conos$new(scon)

# load annotations
ann2 <- readRDS('/d0-mendel/home/meisl/Workplace/BMME/a.data/Selected_Joint_embdding/cellAno/BMME_cell_ano_0130.rds')
typef <- as.factor(ann2)

```

Initiate the analysis, how much of this is important?
We start with some annotation changing and palette thingies
```{r}
  fraction.palette <- c(Benign='#064196',Noninvolved='#2a9606',Involved='#969606',Tumor='#960d06')

  to.coarse.ann <- function(ann) {
    ca <- setNames(as.character(ann),names(ann))
    ca[grep("^CTL",ca)] <- "CD8+ T"
    ca[grep("T regs",ca)] <- "Regulatory T"
    ca[grep("TAM-|Macro",ca)] <- "Macrophage"
    ca[grep("Mono-",ca)] <- "Monocytes"
    ca[grep("mmature B",ca)] <- "Immature B"
    ca[grep("Memory T",ca)] <- "Memory T"
    ca[grep("Monocyte prog",ca)] <- "Monocyte prog"
    ca[grep("Mature B cells",ca)] <- "Mature B"
    ca[grep("Helper T",ca)] <- "Helper T"
    ca[grep("memBcell",ca)] <- "Memory B"
    ca[grep("PDC",ca)] <- "pDC"
    ca <- gsub(" cell.?$","",ca)
    return(as.factor(ca))
  }
  
  typefc <- to.coarse.ann(typef) # here we define typefc, we had to re-arrange the code; it's annot transformed a bit
  
  type.palette <- setNames(rainbow(length(levels(typefc))),levels(typefc))
  pie(1:length(type.palette),labels=names(type.palette),col=type.palette)

  # correct fraction names, return a factor
  get.fractions <- function(x,correct=T) {
    xn <- names(x);
    xf <- gsub(".*-","",as.character(x));
    xf <- gsub("Whole","Benign",xf);
    xf <- factor(xf,levels=c('Benign','Noninvolved','Involved','Tumor'))
    if(!is.null(xn)) names(xf) <- xn;
    xf
  }
  get.patients <- function(x) { gsub("-.*","",x) }

  samplef <- lapply(scon$samples,function(x) rownames(x$counts))
  samplef <- as.factor(setNames(rep(names(samplef),unlist(lapply(samplef,length))),unlist(samplef)))
  fractionf <- get.fractions(samplef)


  samplef <- samplef[names(typef)] # categories but with sample info
  fractionf <- fractionf[names(typef)] # categories
```

Comment 5
fractional changes
```{r}
  # how about fractional changes?
  xt <- table(typefc,samplef)
  xf <- t(t(xt)/colSums(xt))
  df <- melt(xf) # melt is basically gather
  df$patient <- gsub("-.*","",df$samplef); df$fraction <- gsub(".*-","",df$samplef);
  df$samplef <- NULL # to remove the sample-tissue-type col

  # probably should jsut compare between cancer fractions and Bening controls
  df <- melt(xf)
  df$patient <- gsub("-.*","",df$samplef); df$fraction <- gsub(".*-","",df$samplef);
  df$fraction <- get.fractions(df$samplef)
  df <- df %>% arrange(typefc,fraction) # 21 subtypes 32 patient/tissue-cat -> 672 vals

  head(df)
  
  x <- as_tibble(df) %>% group_by(typefc,fraction) %>% summarise_at(vars(value),funs(mean(.,na.rm=T))) %>% ungroup() %>% spread(fraction,value) %>% mutate(diff=Tumor-Benign) %>% arrange(desc(diff)) %>% as.data.frame()
# ^ fraction of cell types in different tissue categories, overall i.e. only 21 rows, used only for subnames it seems
  # sort cell types
  df$typefc <- factor(df$typefc,levels=as.character(x$typefc)) # turning typefc into factor, necessary?

  # full plot (for the supp)
  p <- ggplot(na.omit(df),aes(x=typefc,y=value,dodge=fraction,fill=fraction))+geom_boxplot(notch=FALSE,outlier.shape=NA) +scale_fill_manual(values=fraction.palette) +  geom_point(position = position_jitterdodge(jitter.width=0.1),color=adjustcolor(1,alpha=0.3),aes(pch=fraction),size=0.8)+ theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5))  +xlab("") +ylab("fraction of total cells")+ theme(legend.position="top")
  p # value <- fraction of subtype for a patient-tissue-type combination

  #pdf(file='fraction.all.pdf',width=12,height=6)
  #print(p)
  #dev.off()
```



Weighted distance matrix
```{r}
  sn <- function(x) setNames(x, x)
  cm <- scon$getClusterCountMatrices(groups=typefc)
  cct <- table(typefc,samplef)

  cmpt <- gsub("-.*","",names(cm)) # names of patients
  cmfr <- gsub(".*-","",names(cm)) # tissue type vector

  # without tumor signature correction
  ctdm.nc <- mclapply(setNames(colnames(cm[[1]]), colnames(cm[[1]])),function(ct) {
    tcm <- do.call(rbind,lapply(cm,function(x) x[,ct]))
    tcm <- t(tcm/pmax(1,rowSums(tcm)))
    tcd <- pagoda2:::jsDist(tcm); dimnames(tcd) <- list(colnames(tcm),colnames(tcm));
    # calculate how many cells there are
    attr(tcd,'cc') <- cct[ct,colnames(tcm)]
    tcd
  },mc.cores=1)

  x <- abind(lapply(ctdm.nc,function(x) {
    nc <- attr(x,'cc');
    #wm <- (outer(nc,nc,FUN='pmin'))
    wm <- sqrt(outer(nc,nc,FUN='pmin'))
    return( x*wm )
  }),along=3)
  y <- abind(lapply(ctdm.nc,function(x) {
    nc <- attr(x,'cc');
    sqrt(outer(nc,nc,FUN='pmin'))
  }),along=3)
  xd.nc <- apply(x,c(1,2),sum)/apply(y,c(1,2),sum) # some kind of js distances I guess
  
  require(Rtsne)
  xde <- Rtsne(xd.nc,is_distance=TRUE, perplexity=3,max_iter=1e3)$Y
  df <- data.frame(xde); rownames(df) <- rownames(xd.nc); colnames(df) <- c("x","y");
  df$fraction <- gsub(".*-","",rownames(df))
  df$patient <- gsub("-.*","",rownames(df))
  df$ncells <- colSums(cct)[rownames(df)]

  df$patient <- gsub("BMM.*","Benign",df$patient) 
  df$patient <- gsub("^[A-Z]$","Normal",df$patient) # don't seem to work

  ts.sample.tsne.nc <- ggplot(df,aes(x,y,color=patient,shape=fraction,size=log10(ncells))) + geom_point() + theme_bw() + xlab("") + ylab("") + theme(axis.title=element_blank(),  axis.text=element_blank(), axis.ticks=element_blank()) + guides(color=guide_legend(ncol=2));
  ts.sample.tsne.nc # why are we plotting tsne of js distances?
```

Comment 2
Should be tumor corrected in cell below but let's disregard the correction
```{r}
  ctdm <- ctdm.nc
  x <- abind(lapply(ctdm,function(x) { # line 1409, keep ctdm in cleaned rmd because we need it
    nc <- attr(x,'cc');
    #wm <- (outer(nc,nc,FUN='pmin'))
    wm <- sqrt(outer(nc,nc,FUN='pmin'))
    return( x*wm )
  }),along=3)
  y <- abind(lapply(ctdm,function(x) {
    nc <- attr(x,'cc');
    sqrt(outer(nc,nc,FUN='pmin'))
  }),along=3)

  xd <- apply(x,c(1,2),sum)/apply(y,c(1,2),sum) # first weighting by min count then dividing by all the values to 're-normalize' I guess
  
  xde <- Rtsne(xd,is_distance=TRUE, perplexity=3,max_iter=1e3)$Y
  df <- data.frame(xde); rownames(df) <- rownames(xd); colnames(df) <- c("x","y");
  df$fraction <- gsub(".*-","",rownames(df))
  df$patient <- gsub("-.*","",rownames(df))
  df$ncells <- colSums(cct)[rownames(df)]
  
  df$patient <- gsub("BMM.*","Benign",df$patient)
  df$patient <- gsub("^[A-Z]$","Normal",df$patient)

  ts.sample.tsne <- ggplot(df,aes(x,y,color=patient,shape=fraction,size=log10(ncells))) + geom_point() + theme_bw() + xlab("") + ylab("") + theme(axis.title=element_blank(),  axis.text=element_blank(), axis.ticks=element_blank()) + guides(color=guide_legend(ncol=2));
  ts.sample.tsne

```


distance magnitude comparisons
```{r}
  # distance magnitude comparisons
  # first, on combined distance matrix

  x <- xd.nc; x[upper.tri(x)] <- NA; diag(x) <- NA;
  df2 <- na.omit(melt(x)) # getting distance values

  df2$patient1 <- gsub("-.*","",df2$Var1)
  df2$patient2 <- gsub("-.*","",df2$Var2)
  df2$fraction1 <- gsub(".*-","",df2$Var1)
  df2$fraction2 <- gsub(".*-","",df2$Var2)

  df2$samePatient <- df2$patient1==df2$patient2;
  df2$sameFraction <- df2$fraction1==df2$fraction2;

  df2$withTumor <- df2$fraction1=='Tumor' | df2$fraction2=='Tumor'
  df2$withInvolved <- df2$fraction1=='Involved' | df2$fraction2=='Involved'
  df2$withNoninvolved <- df2$fraction1=='Noninvolved' | df2$fraction2=='Noninvolved'
  
  df2$type <- NA
  df2$type[df2$sameFraction & df2$fraction1=='Whole'] <- 'Benign'
  df2$type[df2$sameFraction & df2$fraction1=='Involved'] <- 'Involved'
  df2$type[df2$sameFraction & df2$fraction1=='Noninvolved'] <- 'Noninvolved'
  df2$type[df2$sameFraction & df2$fraction1=='Tumor'] <- 'Tumor'
  df2$type[df2$sameFraction & df2$fraction1=='Healthy'] <- 'Healthy'
  df2$type <- factor(df2$type,levels=c('Benign','Noninvolved','Involved','Tumor'))

  head(na.omit(df2))
  df2 <- df2[df2$sameFraction==TRUE,]
  ts.within.fraction <- ggplot(na.omit(df2),aes(x=type,y=value))+geom_boxplot(notch=TRUE,outlier.shape=NA,aes(fill=type))+scale_fill_manual(values=fraction.palette)+geom_jitter(position=position_jitter(0.2),color=adjustcolor('black',alpha=0.2))+ theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5)) +  guides(fill=FALSE) + xlab('') + ylab('inter-patient expression distance')
  ts.within.fraction
```

Comment 4
inter-patient distances per cell type
```{r}
  # same thing, but per cell type
  min.cells <- 10
  xl <- do.call(rbind,lapply(names(ctdm.nc),function(xn) {
    x <- ctdm.nc[[xn]] # jsdists for subtype xn
    nc <- attr(x,'cc'); # count of cells
    wm <- outer(nc,nc,FUN='pmin') # min cells for the comparison

    x[upper.tri(x)] <- NA; diag(x) <- NA;
    wm[upper.tri(wm)] <- NA; diag(wm) <- NA;
    df2 <- melt(x);
    df2$ncells <- melt(wm)$value
    df2 <- na.omit(df2)
    df2 <- df2[df2$ncells>=min.cells,];

    df2$patient1 <- gsub("-.*","",df2$Var1)
    df2$patient2 <- gsub("-.*","",df2$Var2)
    df2$fraction1 <- gsub(".*-","",df2$Var1)
    df2$fraction2 <- gsub(".*-","",df2$Var2)

    df2$samePatient <- df2$patient1==df2$patient2;
    df2$sameFraction <- df2$fraction1==df2$fraction2;

    df2$withTumor <- df2$fraction1=='Tumor' | df2$fraction2=='Tumor'
    df2$withInvolved <- df2$fraction1=='Involved' | df2$fraction2=='Involved'
    df2$withNoninvolved <- df2$fraction1=='Noninvolved' | df2$fraction2=='Noninvolved'

    df2$type <- NA
    df2$type[df2$sameFraction & df2$fraction1=='Whole'] <- 'Benign'
    df2$type[df2$sameFraction & df2$fraction1=='Involved'] <- 'Involved'
    df2$type[df2$sameFraction & df2$fraction1=='Noninvolved'] <- 'Noninvolved'
    df2$type[df2$sameFraction & df2$fraction1=='Tumor'] <- 'Tumor'
    df2$type[df2$sameFraction & df2$fraction1=='Healthy'] <- 'Healthy'
    df2$type <- factor(df2$type,levels=c('Benign','Noninvolved','Involved','Tumor'))
    df2 <- df2[df2$sameFraction==TRUE,]
    df2$cell <- xn;
    df2
    
  }))
  
    # leave only cell types with all fractions
  vt <- tapply(xl$fraction1,as.factor(xl$cell),function(x) length(unique(x)));
  xl <- xl[xl$cell %in% names(vt)[vt>3],]
  ts.within.fraction.all <- ggplot(na.omit(xl),aes(x=cell,y=value,dodge=type))+geom_boxplot(notch=TRUE,outlier.shape=NA,aes(fill=type)) +scale_fill_manual(values=fraction.palette) +  geom_point(position = position_jitterdodge(jitter.width=0.1),color=adjustcolor(1,alpha=0.3),aes(pch=type),size=0.8)+ theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5))  +xlab("") +ylab("inter-patient expression distance")+ theme(legend.position="top")
  ts.within.fraction.all
```

Between fractions
```{r}
# now comapre distances between fractions
  # with correction
  x <- xd; x[upper.tri(x)] <- NA; diag(x) <- NA;
  df2 <- na.omit(melt(x))

  df2$patient1 <- gsub("-.*","",df2$Var1)
  df2$patient2 <- gsub("-.*","",df2$Var2)
  df2$fraction1 <- gsub(".*-","",df2$Var1)
  df2$fraction2 <- gsub(".*-","",df2$Var2)

  df2$samePatient <- df2$patient1==df2$patient2;
  df2$sameFraction <- df2$fraction1==df2$fraction2;

  df2$withTumor <- df2$fraction1=='Tumor' | df2$fraction2=='Tumor'
  df2$withInvolved <- df2$fraction1=='Involved' | df2$fraction2=='Involved'
  df2$withNoninvolved <- df2$fraction1=='Noninvolved' | df2$fraction2=='Noninvolved'

  df2$type <- NA
  df2$type[df2$samePatient & df2$withInvolved & df2$withNoninvolved] <- 'Involved vs. Noninvolved'
  df2$type[df2$samePatient & df2$withInvolved & df2$withTumor] <- 'Involved vs. Tumor'
  df2$type[df2$samePatient & df2$withNoninvolved & df2$withTumor] <- 'Noninvolved vs. Tumor'

  ts.fractions <- ggplot(na.omit(df2),aes(x=as.factor(type),y=value))+geom_boxplot(notch=F,outlier.shape=NA)+geom_jitter(position=position_jitter(0.2),aes(color=patient1))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("") +ylab("expression distance")
  ts.fractions
```

Comment 3
Expression shift magnitude thingy
```{r}
  # for a given comparison, look at the magnitude of the differences observed for different cell types
  comp <- c("Noninvolved","Involved")
  comp <- c("Noninvolved","Tumor")
  comp <- c("Involved","Tumor")
  comp <- c("nvolved","Tumor")
  min.cells <- 10
  x <- lapply(ctdm,function(xm) {
    nc <- attr(xm,'cc');
    wm <- outer(nc,nc,FUN='pmin')
    frm <- outer(grepl(comp[1],colnames(xm)),grepl(comp[2],rownames(xm)))==1; # appropriate fractions
    sp <- outer(gsub("-.*","",colnames(xm)),gsub("-.*","",rownames(xm)),FUN='=='); # same patient
    # restrict
    xm[!sp] <- NA;
    xm[!frm] <- NA;
    xm[wm<min.cells] <- NA;
    if(!any(!is.na(xm))) return(NULL);
    xmd <- na.omit(melt(xm))
    wm[is.na(xm)] <- NA;
    xmd$n <- na.omit(melt(wm))$value
    return(xmd);
  })

  x <- x[!unlist(lapply(x,is.null))]
  df <- do.call(rbind,lapply(sn(names(x)),function(n) { z <- x[[n]]; z$cell <- n; z }))
  df$patient <- gsub("-.*","",df$Var1)
  
    # sort cell types
  df$cell <- factor(df$cell,levels=names(sort(tapply(df$value,as.factor(df$cell),median))))

  ts.tumor <- ggplot(na.omit(df),aes(x=as.factor(cell),y=value))+geom_boxplot(notch=FALSE,outlier.shape=NA)+geom_jitter(position=position_jitter(0.1),aes(color=patient))+ theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5))  +xlab("") +ylab("expression distance")+ theme(legend.position="top")
  ts.tumor <- ggplot(na.omit(df),aes(x=as.factor(cell),y=value))+geom_boxplot(notch=FALSE,outlier.shape=NA)+geom_jitter(position=position_jitter(0.1),aes(color=patient),show.legend=FALSE,alpha=0.5)+ theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5))  +xlab("") +ylab("within-patient expression distance")
  ts.tumor
```

