---
title: "jensen_shannon_stuff"
author: "githubz0r"
date: "2019-05-17"
output: workflowr::wflow_html
---

## Introduction

```{r}
library(conos)
library(pheatmap)
#library(fuck)
source('/home/larsc/SecretUtils/R/asdf.R')
source('/home/larsc/SecretUtils/R/peter_code_utils.R')
require(pagoda2)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(irlba)
con_object <- readRDS(file.path('/home/larsc/data/10x_preproced_graphed.rds'))
annot <- readRDS(file.path('/home/demharters/R/projects/UPF9_14_17_19_22_23_24_32_33/metadata_10x_final.rds'))
```

rowbind the adjusted expression values 
```{r}
if (is.null(annot$cellid)) {
  annot$cellid <- annot %>% rownames
}
annot <- annot %>% mutate(subtype_condition = paste(annot$subtype, annot$condition, sep='_'))
rbound_panel <- RbindPanel(con_object)
```

Are the cell names in the same order?
```{r}
identical(rownames(rbound_panel), annot$cellid) # thank god, maybe add a sort for the future in Panelize
```

condition subtype table
```{r}
state_split <- split(annot, annot$condition, drop=TRUE)
condition_tables <- state_split %>% lapply(function(x){table(x$subtype)})
condition_tables
```

plot joint graph colored by sample
```{r}
con_object$plotGraph(color.by='sample',mark.groups=F,alpha=0.1,show.legend=T)
```

subtypes
```{r}
annot <- annot %>% mutate(subtype_condition = paste(annot$subtype, annot$condition, sep='_'))
subannot=setNames(annot$subtype, annot$cellid)
con_object$plotGraph(groups=subannot, font.size=3, shuffle.colors=T, show.legend=F)
```

condition
```{r}
disannot<-setNames(annot$condition, annot$cellid)
con_object$plotGraph(groups=disannot)
```


subtype-condition
```{r}
annot <- annot %>% mutate(subtype_condition = paste(annot$subtype, annot$condition, sep='_'))
cellannot=setNames(annot$subtype_condition, annot$cellid)
con_object$plotGraph(groups=cellannot, font.size=3, shuffle.colors=T, show.legend=F)
```

Individual subtype joint graph plot
```{r}
plotOneSubtype <- function(con.object, annotation, subtype, font.size=2, alpha=0.3, size=0.4){
  split.annot<-split(annotation, annotation$subtype)
  sub.annot <- split.annot[[subtype]]
  sub.annot <- sub.annot %>% mutate(sub.cond = paste(sub.annot$subtype, sub.annot$condition, sep='_'))
  sub.groups <- setNames(sub.annot$sub.cond, sub.annot$cellid)
  con.object$plotGraph(groups=sub.groups, font.size=font.size, alpha=alpha, size=size, mark.groups=T, plot.na=F)
}
#plotOneSubtype(con_object, annot, 'L2_Lamp5') # really should use repel, but I can't make it work
all_types <- annot$subtype %>% unique
all_types_plots <- all_types %>% lapply(function(x, con.obj, annotation){
  plotOneSubtype(con.obj, annotation, x, font.size=3, alpha=0.5, size=1.5)}, con_object, annot)
plot_grid(plotlist=all_types_plots[1:4], nrow=2)
plot_grid(plotlist=all_types_plots[5:8], nrow=2)
plot_grid(plotlist=all_types_plots[9:12], nrow=2)
plot_grid(plotlist=all_types_plots[13:16], nrow=2)
plot_grid(plotlist=all_types_plots[17:20], nrow=2)
```


```{r}
plotOneSubtype(con_object, annot, 'L4_Rorb', font.size=3, alpha=0.5, size=1.5)
```

fractional plot
```{r}
FractionalPlot(annot$sample, annot$subtype, annot$condition)
```

More details on individual samples
```{r, fig.height=12}
frac_df <- FractionalPlot(annot$sample, annot$subtype, annot$condition, return.plot=F)
ggplot(na.omit(frac_df),aes(x=subtype,y=freq))+geom_bar(stat='identity')+
    theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5)) +
    xlab("") +ylab("fraction of total cells")+facet_wrap(~patient, nrow=4)
```

```{r}
ggplot(na.omit(frac_df),aes(x=subtype,y=freq, col=patient, shape=condition))+geom_point()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 90, hjust = 0.5)) +
    xlab("") +ylab("fraction of total cells")
```

PCA
```{r}
od_genes = conos:::getOdGenesUniformly(con_object$samples, 4000)
pca_cm <- prcomp_irlba(rbound_panel[, od_genes],n=100)
pca_cmat <- pca_cm$x
rownames(pca_cmat) <- rownames(rbound_panel)
```
plot PCA eigenspectrum
```{r}
pca_sum <- summary(pca_cm)
bind_cols(percent_var=pca_sum$importance[2,], number=c(1:100)) %>% ggplot(aes(y=percent_var, x=number))+geom_point()
```

PCA annotated by samples
```{r}
sampannot <- setNames(annot$sample, annot$cellid)
pca_cmat[,1:2] %>% as_tibble %>%  mutate(samples=annot$sample) %>% 
  ggplot(aes(x=PC1, y=PC2))+geom_point(aes(col=samples), alpha=0.3, size=0.2)+guides(colour = guide_legend(override.aes = list(size=2, alpha=1)))
```

Tsne, rbound pagoda looks like sample 4 has a lot of Lamp5 cells
```{r}
require(Rtsne)

pagoda_tsne <- pca_cmat[,0:25] %>% as.matrix %>% Rtsne(pca=F)
tsne_vals <- pagoda_tsne$Y; colnames(tsne_vals)=c('var1', 'var2')
tsne_annot <- bind_cols(tsne_vals %>% as.data.frame, annot)
tsne_annot %>% ggplot(aes(x=var1, y=var2))+geom_point(aes(col=sample), alpha=0.7, size=0.2) + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```


Tsne whole pagoda
```{r}
devtools::load_all('/home/viktor_petukhov/Copenhagen/NeuronalMaturation')
cm_merged_raw <- lapply(con_object$samples, function(p2) t(p2$misc$rawCounts)) %>%
  NeuronalMaturation::MergeCountMatrices()

p2 <- NeuronalMaturation::GetPagoda(cm_merged_raw)

```

pagoda PCA
```{r}
pca_whole <- bind_cols(pca1=p2$reductions$PCA[,1], pca2=p2$reductions$PCA[,2], sample=annot$sample, 
                       subtype=annot$subtype, condition=annot$condition)

pca_whole %>% ggplot(aes(x=pca1, y=pca2))+geom_point(aes(col=sample), alpha=0.3, size=0.2)+guides(colour = guide_legend(override.aes = list(size=2, alpha=1)))
```

```{r}
conos:::plotSamples(list(p2), groups=sampannot, shuffle.colors=T, font.size=c(2,5),
                    show.legend=T, size=0.4)
```

```{r}
MergedOneSubtype <- function(annotation, subtype, font.size=2, alpha=0.3, size=0.4){
  split.annot<-split(annotation, annotation$subtype)
  sub.annot <- split.annot[[subtype]]
  sub.annot <- sub.annot %>% mutate(sub.cond = paste(sub.annot$subtype, sub.annot$sample, sep='_'))
  sub.groups <- setNames(sub.annot$sub.cond, sub.annot$cellid)
  conos:::plotSamples(list(p2), groups=sub.groups, shuffle.colors=T, font.size=c(2),
                    show.legend=T, size=0.4, plot.na=F)
}
#plotOneSubtype(con_object, annot, 'L2_Lamp5') # really should use repel, but I can't make it work
all_types <- annot$subtype %>% unique
all_types_plots <- all_types %>% lapply(function(x, annotation){MergedOneSubtype(annotation, x)}, annot)
#names(all_types_plots) <- all_types
all_types_plots[10]
```


```{r}
#plot_grid(all_types_plots)
```

